---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Download the data.
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, "repdata_data_activity.zip", method="curl")
```
### Unzip the data. 
```{r}
unzip("repdata_data_activity.zip")
```
### Read in the data
```{r}
activity_data <- read.csv("activity.csv")
```
### Convert the "date" variable to date class.
```{r}
activity_data$date <- as.Date(activity_data$date)
```

## What is mean total number of steps taken per day?
### Create a histogram of steps by day. 
```{r}
steps_by_day <- tapply(activity_data$steps, activity_data$date, FUN=sum, na.rm=TRUE)
hist(steps_by_day)
```

## Calculate the means for each day.
```{r}
tapply(activity_data$steps, activity_data$date, FUN=mean, na.rm=TRUE)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
