---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
#### Download the data.
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, "repdata_data_activity.zip", method="curl")
```
#### Unzip the data. 
```{r}
unzip("repdata_data_activity.zip")
```
#### Read in the data
```{r}
activity_data <- read.csv("activity.csv")
```
#### Convert the "date" variable to date class.
```{r}
activity_data$date <- as.Date(activity_data$date)
```

## What is mean total number of steps taken per day?
#### Create a histogram of steps by day. 
```{r}
steps_by_day <- tapply(activity_data$steps, activity_data$date, FUN=sum, na.rm=TRUE)
hist(steps_by_day, main = "Steps per day", xlab = "Number of Steps", ylab= "Number of Days")
```

#### Calculate the mean and median steps per day. 
``` {r echo = FALSE}
steps_mean <- mean(steps_by_day)
steps_median <- median(steps_by_day)
```
The mean of steps each day is `r steps_mean`. 
The median of steps each day is `r steps_median`.

## What is the average daily activity pattern?
#### Create a time series plot.
```{r}
steps_by_interval <- tapply(activity_data$steps, activity_data$interval, FUN=mean, na.rm = TRUE)
plot(steps_by_interval, 
     type="l", 
     main="Steps during each interval", 
     xlab = "Interval", 
     ylab = "Average steps taken")
```

```{r echo=FALSE}
max_index <- activity_data$interval[which.max(steps_by_interval)]
```
The interval with the maximum number of steps
is `r max_index`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
